---
title: "midterm project"
author: "Hao He"
date: "2022-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(readxl)
```

# Read in data and pre-processing

## Examine the columns 
```{r strawberries}
strawb <- read_xlsx("strawberries-2022oct30-a.xlsx", col_names = TRUE)

# Get the column names and index them
cnames <- colnames(strawb)
x <- 1:dim(strawb)[2]

```

Create an indicator T to check the unique values of each one of the columns, and then use T to select columns to drop
```{r}
T <- NULL
for(i in x){T <- c(T, dim(unique(strawb[i]))[1])}
drop_cols <- cnames[which(T == 1)]
strawb %<>% select(!all_of(drop_cols))

# Arrange columns by year and state
strawb %<>% arrange(Year, State)

```


Now look at the `Data Item` column, it has a couple columns packed into one. We should use separate() to split it up into 3 or 4 columns.
```{r}
# colnames(strawb)
temp1 <- strawb %>% select(`Data Item`) %>% 
         distinct()
# 3 columns
strawb2 <- strawb %>% separate(col=`Data Item`,
                into = c("Strawberries", "items", "units"),
                sep = ",",
                fill = "right")

# Warning says discarding additional info, so we need 4 columns
strawb3 <- strawb %>% separate(col=`Data Item`,
            into = c("Strawberries", "type", "items", "units"),
                               sep = ",",
                               fill = "right")

# For `strawberries` column, we may have to use separate again to split up the `price received` part in further analysis, if we want to compare to price received data.
temp2 <- strawb3 %>% select(Strawberries) %>% 
  distinct()

```


## Create susbets: organic strawberries, non-organic strawberries, chemicals used in strawberries.
Looking at the `domain category` column, it may be better to break down the strawberry dataset into subsets for further analysis on organic vs non-organic, and commercials vs chemicals used in strawberries.

Let's create the organic and non-organic strawberries subsets first.
```{r}
# type: 62 rows
type_organic <- strawb3 %>% filter(grepl("organic", 
               .$type, 
               ignore.case = T))

# zero rows returned 
items_organic <- strawb3 %>% filter(grepl("organic", 
                      .$items, 
                      ignore.case = T))  
# Domain: 62 rows
Domain_organic <- strawb3 %>% filter(grepl("organic", 
                      .$Domain, 
                      ignore.case = T))

#  Domain Category: 62 rows
Domain_Category_organic <- strawb3 %>% filter(grepl("organic", 
                       .$`Domain Category`, 
                       ignore.case = T))

# combine all the organic entries
(intersect(type_organic, Domain_organic) == intersect(type_organic, Domain_Category_organic))

org_rows <- intersect(type_organic, Domain_organic)

strawb_organic <- left_join(org_rows, strawb3)

strawb_non_organic <- anti_join(strawb3, org_rows)
```


Next, separate the chemical data. We start with non-organic data because they have a lot more info of chemicals than that of organic data.

```{r}
# check what could be indications of chemical data 
(strawb_non_organic %>% select(type) %>% distinct())

# the chemical data in the column "type" is in the rows where type = BEARING - Application, so we find them by matching strings 

chem_rows <- strawb_non_organic%>% filter(grepl("BEARING - APPLICATIONS", 
                      .$type, 
                      ignore.case = T)) 

# 2112 rows matches for chemical data in non-organic subset.

# The column `Domain` and `Domain Category` also contain info about chemicals

chem_rows_D <- strawb_non_organic%>% filter(grepl("chemical", 
                      .$Domain, 
                      ignore.case = T))
chem_rows_D_C <- strawb_non_organic%>% filter(grepl("chemical", 
                      .$`Domain Category`, 
                      ignore.case = T))

ins <- intersect(chem_rows, chem_rows_D)
ins_2 <- intersect(chem_rows, chem_rows_D_C)

# All of the row numbers in chem_rows_D and chem_rows_D_C are are in chem_rows. 
# So it means chem_rows contain all of the chemical data.

```

After we get all rows of chemical data, now we can create the chemicals subset.
```{r}
strawb_chem <- left_join(chem_rows, strawb_non_organic)
```


Next, try to find how many hazardous chemicals mentioned in the "Shoppers Guide to Pesticides in Produce" exists in our dataset. Hazardous chemicals include Carbendazim, Bifenthrin, methyl bromide, 1,3-dichloropropene, chloropicrin, Telone.
```{r}
# carbendazim: 0 matches
df_carbendazim <- strawb3 %>% filter(grepl("Carbendazim", .$`Domain Category`, ignore.case=TRUE))

# Bifenthrin: 27 matches
df_Bifenthrin <- strawb3 %>% filter(grepl("Bifenthrin", .$`Domain Category`, ignore.case=TRUE))

# methyl bromide: 3 matches
df_methyl_bromide <- strawb3 %>% filter(grepl("methyl bromide", .$`Domain Category`, ignore.case=TRUE))

# 1,3-dichloropropene: 0 matches 
df_1_3_dichloropropene <- strawb3 %>% filter(grepl("1,3-dichloropropene", .$`Domain Category`, ignore.case=TRUE))

# chloropicrin: 18 matches
df_chloropicrin <- strawb3 %>% filter(grepl("chloropicrin", .$`Domain Category`, ignore.case=TRUE))

# Telone: 0 matches
df_Telone <- strawb3 %>% filter(grepl("Telone", .$`Domain Category`, ignore.case=TRUE))
```

## Including Plots
