---
title: "midterm project"
author: "Hao He"
date: "2022-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(readxl)
```

# Read in data and pre-processing

## Examine the columns 
```{r strawberries}
strawb <- read_xlsx("strawberries-2022oct30-a.xlsx", col_names = TRUE)

# Get the column names and index them
cnames <- colnames(strawb)
x <- 1:dim(strawb)[2]

```

Create an indicator T to check the unique values of each one of the columns, and then use T to select columns to drop
```{r}
T <- NULL
for(i in x){T <- c(T, dim(unique(strawb[i]))[1])}
drop_cols <- cnames[which(T == 1)]
strawb %<>% select(!all_of(drop_cols))

# Arrange columns by year and state
strawb %<>% arrange(Year, State)

```


Now look at the `Data Item` column, it has a couple columns packed into one. We should use separate() to split it up into 3 or 4 columns.
```{r}
# colnames(strawb)
temp1 <- strawb %>% select(`Data Item`) %>% 
         distinct()
# 3 columns
strawb2 <- strawb %>% separate(col=`Data Item`,
                into = c("Strawberries", "items", "units"),
                sep = ",",
                fill = "right")

# Warning says discarding additional info, so we need 4 columns
strawb3 <- strawb %>% separate(col=`Data Item`,
            into = c("Strawberries", "type", "items", "units"),
                               sep = ",",
                               fill = "right")

# For `strawberries` column, we may have to use separate again to split up the `price received` part in further analysis, if we want to compare to price received data.
temp2 <- strawb3 %>% select(Strawberries) %>% 
  distinct()

```


## Create susbets: organic strawberries, non-organic strawberries, chemicals used in strawberries.
Looking at the `domain category` column, it may be better to break down the strawberry dataset into subsets for further analysis on organic vs non-organic, and commercials vs chemicals used in strawberries.

Let's create the organic and non-organic strawberries subsets first.
```{r}
# type: 62 rows
type_organic <- strawb3 %>% filter(grepl("organic", 
               .$type, 
               ignore.case = T))

# zero rows returned 
items_organic <- strawb3 %>% filter(grepl("organic", 
                      .$items, 
                      ignore.case = T))  
# Domain: 62 rows
Domain_organic <- strawb3 %>% filter(grepl("organic", 
                      .$Domain, 
                      ignore.case = T))

#  Domain Category: 62 rows
Domain_Category_organic <- strawb3 %>% filter(grepl("organic", 
                       .$`Domain Category`, 
                       ignore.case = T))

# combine all the organic entries
(intersect(type_organic, Domain_organic) == intersect(type_organic, Domain_Category_organic))

org_rows <- intersect(type_organic, Domain_organic)

strawb_organic <- left_join(org_rows, strawb3)

strawb_non_organic <- anti_join(strawb3, org_rows)
```


Next, separate the chemical data. We start with non-organic data because they have a lot more info of chemicals than that of organic data.

```{r}
# check what could be indications of chemical data 
(strawb_non_organic %>% select(type) %>% distinct())

# the chemical data in the column "type" is in the rows where type = BEARING - Application, so we find them by matching strings 

chem_rows <- strawb_non_organic%>% filter(grepl("BEARING - APPLICATIONS", 
                      .$type, 
                      ignore.case = T)) 

# 2112 rows matches for chemical data in non-organic subset.

# The column `Domain` and `Domain Category` also contain info about chemicals

chem_rows_D <- strawb_non_organic%>% filter(grepl("chemical", 
                      .$Domain, 
                      ignore.case = T))
chem_rows_D_C <- strawb_non_organic%>% filter(grepl("chemical", 
                      .$`Domain Category`, 
                      ignore.case = T))

ins <- intersect(chem_rows, chem_rows_D)
ins_2 <- intersect(chem_rows, chem_rows_D_C)

# All of the row numbers in chem_rows_D and chem_rows_D_C are are in chem_rows. 
# So it means chem_rows contain all of the chemical data.

```

After we get all rows of chemical data, now we can create the chemicals subset.
```{r}
strawb_chem <- left_join(chem_rows, strawb_non_organic)
```

##  Cleaning three tibbles (subsets)

### Chemical data subset

Write a function to drop columns that contain useless information. 
```{r}
before_cols = colnames(strawb_chem)
  T = NULL
  x = length(before_cols)

   for(i in 1:x){
    b <- length(unlist(strawb_chem[,i] %>% unique()) )
    T <- c(T,b)
   }
    
  drop_cols <- before_cols[which(T == 1)]
  after_strawb_chem <- strawb_chem %>% select(!all_of(drop_cols))
  after_cols <-  colnames(after_strawb_chem)
```

It's easy to find that after we drop some useless columns, the `units` column still contain NA values in the form of character data. Also, we need to separate Domain Category column to get a cleaner data frame.
```{r}
after_strawb_chem %>% select(units) %>% distinct()  # NA or AVG as row values

# separate `Domain Category` column
strawb_chem <- after_strawb_chem %>% separate(col=`Domain Category`, 
                                    into = c("dc1", "chem_name"),
                                    sep = ":", 
                                    fill = "right")
temp3 <- strawb_chem %>% select(chem_name) %>% unique()
# number of unique rows in chem_name
length(unlist(temp3))

# Examine the `items` column. If we remove the "measured in", it actually corresponds to the units. So rename units to Category and rename items to Units, and now items should only contain unit names.

grep("measured in", 
                    strawb_chem$items, 
                    ignore.case = T) %>% length()

strawb_chem$items <- str_remove_all(strawb_chem$items, "MEASURED IN ")

strawb_chem<-  strawb_chem %>% rename(c(Category = units, Units = items))

# check if Domain and dc1 refers to the same thing? True
sum(strawb_chem$Domain == strawb_chem$dc1) == dim(strawb_chem)[1]


# select the column that is useful for further analysis
strawb_chem <- strawb_chem %>% select(Year, State, Category, Units, dc1, chem_name, Value)
```

Now continue examining the `dc1` column. Check if every entry starts with"CHEMICAL", if so, remove it for better view.

```{r}
bb  <- grep("CHEMICAL, ", 
            strawb_chem$dc1, 
            ignore.case = T)
length(bb)
chem <- 1:2112

non_chem_rows <- setdiff(chem, bb)
length(non_chem_rows)

# 45 entries starts with some other strings: FERTILIZER, keep them for now
strawb_chem %>% slice(non_chem_rows)

# Remove repeated "CHEMICAL, " and rename dc1 to Chem_type 
strawb_chem$dc1 <- str_remove_all(strawb_chem$dc1, "CHEMICAL, ")

strawb_chem$dc1 %>% unique()

strawb_chem %>% rename(Chem_types = dc1)

#Now clean the chem_name column. Remove parentheses and split the chem_name into chem_name and chenm_code.

strawb_chem$chem_name <- str_remove_all(strawb_chem$chem_name, "\\(")
strawb_chem$chem_name <- str_remove_all(strawb_chem$chem_name, "\\)")

strawb_chem <- strawb_chem %>% separate(col = chem_name,
                          into = c("Chem_name","Chem_code"),
                          sep = "=",
                          fill = "right"
                            ) 
```

There are still some NAs in the category column, now fill in a label for NA. First, check that "lb" in the `Units` column corresponds to NA in the category column.
```{r}
aa <- which(strawb_chem$Units == " LB")

bb <- which(is.na(strawb_chem$Category))

sum(aa==bb)==length(aa)

# Fill in label "AVG" for NA
strawb_chem <- strawb_chem %>% 
     replace(is.na(strawb_chem), "AVG")
```

The number of different chemicals listed in our dataset is 172. Since California contributes to 80% of the strawberry production and followed by Florida and Oregon, we may want to check how many different chemicals are listed by state and if California use more chemicals than other states?
```{r}
strawb_chem %>% select(Chem_name) %>% distinct()

# number of different chemicals listed by different states
strawb_chem %>% group_by(State) %>% summarise(unqiue_chems = n_distinct(Chem_name))
```
Based on the tibble above, we can say that 23 more chemicals have been used in California than in Florida, and 118 more chemicals have been used in California than that in Oregon.

Next, try to find how many hazardous chemicals mentioned in the "Shoppers Guide to Pesticides in Produce" exists in our dataset. Hazardous chemicals include Carbendazim, Bifenthrin, methyl bromide, 1,3-dichloropropene, chloropicrin, Telone.
Also, check the hazardous chemicals listed by state.
```{r}
# carbendazim: 0 matches
df_carbendazim <- strawb_chem %>% filter(grepl("Carbendazim", .$Chem_name, ignore.case=TRUE))

# Bifenthrin: 27 matches
df_Bifenthrin <- strawb_chem %>% filter(grepl("Bifenthrin", .$Chem_name, ignore.case=TRUE))
df_Bifenthrin %>% group_by(State) %>% count()

# methyl bromide: 3 matches
df_methyl_bromide <- strawb_chem  %>% filter(grepl("methyl bromide", .$Chem_name, ignore.case=TRUE))
df_methyl_bromide %>% group_by(State) %>% count()

# 1,3-dichloropropene: 0 matches 
df_1_3_dichloropropene <- strawb_chem  %>% filter(grepl("1,3-dichloropropene", .$Chem_name, ignore.case=TRUE))

# chloropicrin: 18 matches
df_chloropicrin <- strawb_chem  %>% filter(grepl("chloropicrin", .$Chem_name, ignore.case=TRUE))
df_chloropicrin %>% group_by(State) %>% count()
# Telone: 0 matches
df_Telone <- strawb_chem  %>% filter(grepl("Telone", .$Chem_name, ignore.case=TRUE))
```

Based on my research safe chemicals to use include Neem Oil and Spinosad. Check if they are in our chemical dataset.
```{r}
# Neem Oil: 27 matches
df_Neem_Oil <- strawb_chem %>% filter(grepl("NEEM OIL", .$Chem_name, ignore.case=TRUE))
df_Neem_Oil %>% group_by(State) %>% count()

#Spinosad: 12 matches 
df_Spinosad <- strawb_chem %>% filter(grepl("SPINOSAD", .$Chem_name, ignore.case=TRUE))
df_Spinosad %>% group_by(State) %>% count()
```


### Organic strawberry subset

Starts with organic_strawb to select columns to drop
```{r}
before_cols = colnames(strawb_organic)
  T = NULL
  x = length(before_cols)

   for(i in 1:x){
    b <- length(unlist(strawb_organic[,i] %>% unique()) )
    T <- c(T,b)
   }
    
  drop_cols <- before_cols[which(T == 1)]
  after_strawb_organic <- strawb_organic %>% select(!all_of(drop_cols))
  after_cols <-  colnames(after_strawb_organic)
```


Check all entries that contain "SALES"
```{r}
type_sales <- grep("sales", 
               after_strawb_organic$type, 
               ignore.case = T) %>% length()

items_sales <- grep("sales", 
               after_strawb_organic$items, 
               ignore.case = T) %>% length()

sum(type_sales,items_sales) == dim(after_strawb_organic)[1]


# Looking at these three columns together, we can see every row contain a "SALES".
# It doesn't add any additional info for further analysis. So we can remove them.
after_strawb_organic$type <- str_remove_all(after_strawb_organic$type, "- SALES")

(after_strawb_organic %>% select(items) %>% unique())

(after_strawb_organic %>% select(units) %>% unique())
```

By checking the unique values of the `items` and `units` column, we can found that there are something in common in these two columns. If we combine them together and then only keep units like" $" and "CWT", we can get the units corresponds to strawberries. Now we can get a clean strawb_organic data frame.
```{r}
# Combine items and units into one column called `Units`
after_strawb_organic <- after_strawb_organic %>% unite(Units,items:units, remove = TRUE)

# Remove "MEASURED IN", "FRESH MARKET - SALES_" and "PROCESSING - SALES_" and "_NA"
after_strawb_organic$Units <- str_remove_all(after_strawb_organic$Units, "MEASURED IN ")
after_strawb_organic$Units <- str_remove_all(after_strawb_organic$Units, "FRESH MARKET - SALES_")
after_strawb_organic$Units <- str_remove_all(after_strawb_organic$Units, "PROCESSING - SALES_")
after_strawb_organic$Units <- str_remove_all(after_strawb_organic$Units, "_NA")


# Drop `State ANSI` column and select columns for further analysis
strawb_organic <- after_strawb_organic %>% select(-c(`State ANSI`,type))

```

### non-organic strawberry subset

Similar cleaning approaches for chemicals can be applied to clean non-organic strawberry subset.
```{r}
# Examine columns
before_cols = colnames(strawb_non_organic)
  T = NULL
  x = length(before_cols)

   for(i in 1:x){
    b <- length(unlist(strawb_non_organic[,i] %>% unique()) )
    T <- c(T,b)
   }
    
  drop_cols <- before_cols[which(T == 1)]
  after_strawb_non_organic <- strawb_non_organic %>% select(!all_of(drop_cols))
  after_cols <-  colnames(after_strawb_non_organic)
  
after_strawb_non_organic %>% select(units) %>% distinct()  # NA or AVG as row values

# separate `Domain Category` column
strawb_non_organic <- after_strawb_non_organic %>% separate(col=`Domain Category`, 
                                    into = c("dc1", "chem_name"),
                                    sep = ":", 
                                    fill = "right")

temp4 <- strawb_non_organic %>% select(chem_name) %>% distinct()

# number of unique rows in chem_name
length(unlist(temp4))

# Examine the `items` column. Check unique row values in `items` column
strawb_non_organic %>% select(items) %>% distinct()

# If we remove the "measured in", it actually corresponds to the units. So rename units to Category and rename items to Units, and now items should only contain unit names.
grep("measured in", 
                    strawb_non_organic$items, 
                    ignore.case = T) %>% length()


strawb_non_organic$items <- str_remove_all(strawb_non_organic$items, "MEASURED IN ")

strawb_non_organic<-  strawb_non_organic %>% rename(c(Category = units, Units = items))

# check if Domain and dc1 refers to the same thing? FALSE
sum(strawb_non_organic$Domain == strawb_non_organic$dc1) == dim(strawb_non_organic)[1]

# now check how many rows of the `Domain` and `dc1` column are not the same: 60 rows
temp5 <- strawb_non_organic[which(strawb_non_organic$Domain != strawb_non_organic$dc1),]
dim(temp5)[1]

# Need to keep these 60 rows because they have value for `Value` column

# Select the column that is useful for further analysis
strawb_non_organic <- strawb_non_organic %>% select(Year, State, Category, Units, Domain, dc1, chem_name, Value)
```




# Answers to Gradescope Question

## Q2 Confidence Interval
Compute a 95% confidence interval for California organic strawberry sales in 2016.
```{r}
strawb_organic$Value <- as.integer(strawb_organic$Value)
table1 <- strawb_organic %>% filter(Year == "2016" & State == "CALIFORNIA") 

# resale the CWT 
table1$Value[2] <- table1$Value[2]*100
table1$Value[4] <- table1$Value[4]*100
table1$Value[6] <- table1$Value[6]*100

n <- dim(table1)[1]
xbar <- mean(table1$Value)
s <- sd(table1$Value)

margin <- qt(0.975,df=n-1)*s/sqrt(n)

CI_95_organic <- c(lowerinterval = xbar - margin, upperinterval = xbar + margin)
CI_95_organic
```

## Q3 Confidence Interval 
Compute a 95% confidence interval for California non-organic strawberry sales in 2016.
```{r}

```

## Q4 Chem Count - CA
Q: In the data set for the MA615 Strawberry project, how many different chemicals are listed? 
A: 139 in CALIFORNIA.
```{r}
strawb_chem %>% group_by(State) %>% summarise(unqiue_chems = n_distinct(Chem_name))
```

## Q5  Chem Count
Q: On the basis of the data set for the MA615 Strawberry project, how many more chemicals have been used in California than in Florida?
A: 23 more in CALIFORNIA than in Florida based on the tibble above.